Quantitative neuroimaging with ANTs

### 2018-02-15 antsr/antsr.Rmd
ENH:  Final commit.
ENH:  ASNR 2017
ENH:  Add Cohen presentation.

# Current work and Advanced Normalization Tools in R (ANTsR)

@@

![](./competitions/figures/brats2013results2.png)

@@

![](./wmhs/figures/wmhPipeline.png)

@@

![](./antsr/figures/filiano_rsfmri.png)

@@

* Corrective learning for segmentation refinement

>>>>

# Advanced Normalization Tools in R (ANTsR)

@@

\begin{centering}

\includegraphics[width=4.5in]{./competitions/figures/brats2013results1.png}

\end{centering}

\small

@@

\begin{centering}

\includegraphics[width=4.5in]{./wmhs/figures/wmhPipeline.png}

\end{centering}

@@

\begin{centering}

\includegraphics[width=3in]{./antsr/figures/filiano_rsfmri.png}

\end{centering}

@@

* [ANTsRNet](https://github.com/stnava/ANTsRNet): a growing collection of popular deep learning architectures

### 2018-02-15 cohen.Rmd
ENH:  Add Cohen presentation.

---
title: \`\`Dr. Tustison (UVA) presentation\'\'
subtitle:
author: Nick Tustison
bibliography: REFERENCES.bib
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme:  professionalfonts
    highlight: tango
    incremental: false
    theme:  AnnArbor
    includes:
      in_header: mystyle.tex
 #   toc: yes
    keep_tex: yes
    slide_level: 2
---

# Developers and collaborators

##  Founders:  Brian and Nick

![](./figures/brian_and_nick.png)

<!--
NOtes:
* Worked together for almost 10 years
* have learned tons
* have written quite a bit of code together
-->

![](./figures/antsCollaborators2.png)

@@

<!--
NOtes:
* However, it is not just us
* Other people have done quite a bit
* and others who are not shown such as Arman
-->
# ANTs lineage

## Image mapping and perception: 1877

Francis Galton: _Can we see criminality in the face?_

![](./figures/galton.png)

_What about syphilis, mental illness?_

<!--
Notes:
* Use of "composite portraiture" and what he called "eugenics"
* Can we see criminality, mental illness, syphilis, and others in facial features?
* Despite the obvious prejudices driving his research, he was a pioneer.
-->

@@

![](./background/figures/FrancisGalton2.jpg)
-->

<!--
Notes:
* Another example where he has manually categorized facial features and try to combine them
* Did not find much---composite tends to blend away to normality.
-->

##  Speaking of criminality...

_Can we say anything about the U.S. Congress?_

![](./background/figures/congressPersons.png)

__Maybe they should have used [ANTs](https://github.com/ntustison/CongressionalFaceTemplates)?__

<!--
Notes:
-->

##  Image mapping & biology:  1917

D'Arcy Thompson: _Comparison of related forms_

![](./figures/dthompson.png)
-->

<!--
Notes:
* Book: On Growth and Form
* Descriptive
* "This book of mine has little need of preface, for indeed it is 'all preface' from beginning to end."
-->

## Image mapping & biology: Current
<BR><BR>
_Look where we are now!_
<BR><BR>
![](./background/figures/r16xr85.png)

<!--
## ANTs family tree

![](./figures/lineage.jpg)
-->

<!--
References: @Horn1981, @Gee1993, @Grenander1993, @Thompson2001, @Miller2002, @Shen2002, @Arnold2014, @Thirion1998, @Rueckert1999, @Fischl2012, @Ashburner2012
-->

<!--
## Initial scope

![](./figures/ants_initial_scope.png)
-->

<!--
Notes:
* Brian work with Murray Grossman
* First registration paper
  - WBIR submission
  - FEM registraiton using B-spline basis functions.
  - "yeah, that registration is not that good---Jim would not like it."
-->

@@

<BR><BR>
_\`\`Papers are just advertisements for the science.\'\'_
<BR>

@@

NOtes:
* Previously discussed Brians work
* The variant most widely used
-->

@@

     -i, --initialize-transforms-per-stage (1)/0
          Initialize linear transforms from the previous stage. By enabling this option,
          the current linear stage transform is directly intialized from the previous
          stage's linear transform; this allows multiple linear stages to be run where
          each stage directly updates the estimated linear transform from the previous
          stage. (e.g. Translation -> Rigid -> Affine).
          <VALUES>: 0

@@

## Atropos + N4 $\rightarrow$ `antsAtroposN4.sh`

@@

## Multi-atlas segmentation

@@

![](./evaluation/figures/templates.png)

<!--
* Templates representing the average mean shape and intensity are built directly from the cohort to be analyzed, e.g. pediatric vs. middle-aged brains.
* Acquisition and anonymization (e.g. defacing) protocols are often different.
-->

@@

![](./evaluation/figures/templatePriors.png)

<!--
Each template is [processed](https://github.com/ntustison/antsCookTemplatePriorsExample)
to produce auxiliary images which are used for brain extraction and brain segmentation.
-->

@@

![](./evaluation/figures/brainExtraction.png)

Comparison with de facto standard FreeSurfer package.  Note the difference in separation of
the gray matter from the surrounding CSF.  (0 failures out of 1205 scans)
-->

@@

![](./evaluation/figures/brainSegmentation.png)

Randomly selected healthy individuals.  ```Atropos``` gets good performance across ages.
-->

@@

![](./evaluation/figures/corticalThicknessEstimation.png)

@@

<!--
##  What about using both the T1 & T2?
-->

@@

<!--
----

@@

# Joint label/intensity fusion

## Multi-atlas labeling

## Multi-atlas segmentation

## Possible uses

* ``Correct'' images
    * motion correction
    * lesions
* Project atlas set intensity signature
-->

>>>>

---
title: Quantitative neuroimaging with ANTs
subtitle:
author: Nick Tustison
bibliography: REFERENCES.bib
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme:  professionalfonts
    highlight: tango
    incremental: false
    theme:  AnnArbor
    includes:
      in_header: mystyle.tex
 #   toc: yes
    keep_tex: yes
    slide_level: 2
    fig_width: 6
    fig_height: 4.5
    fig_caption: false
---

---

# What is ANTs?

## NLM's The Visible Human Project and The Insight Toolkit

\begin{center}

\includegraphics[width=4.75in]{./figures/VisibleHumanProject_ITK.pdf}

\end{center}

## ANTs _began with a clinical research need._

\begin{center}

\includegraphics[width=4.75in]{./figures/ANTsEvolutionarySummary.pdf}

\end{center}

## Founding developers

\begin{center}
\includegraphics[width=4.75in]{./figures/brian_and_nick.png}
\end{center}

\begin{center}
BBA \& NJT
\end{center}

## Collaborators

\begin{center}

\includegraphics[width=4.75in]{./figures/antsCollaborators2.png}

\end{center}

@@

## Advanced Normalization Tools

\begin{center}

\includegraphics[width=4.75in]{./tools/figures/coreANtsToolsNeuro.png}

\end{center}

## Advanced Normalization Tools $\rightarrow$ ``ITK-Lung''

\begin{center}

\includegraphics[width=4.75in]{./tools/figures/coreANtsToolsLung.png}

\end{center}

@@

# Competitions

## Algorithmic fight club

![](./competitions/figures/fightclub.jpg)

## International competitions

* [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496): MRI brain registration

* [EMPIRE 2010](http://empire10.isi.uu.nl):  CT lung registration

* [Multi-Atlas Label Challenge 2012](https://masi.vuse.vanderbilt.edu/workshop2012/index.php/Main_Page):  MRI brain registration and segmentation

* [SATA Challenge 2013](https://masi.vuse.vanderbilt.edu/workshop2013/index.php/MICCAI_2013_SATA_Challenge_and_Workshop:Current_events): MRI cardiac and canine hind leg registration

* [BRATS 2013](http://martinos.org/qtim/miccai2013/): Multi-modal MRI brain segmentation

* [STACOM 2014 MoCo Challenge](http://www.cardiacatlas.org/web/stacom2014/moco-introduction):  MRI cardiac motion estimation

<!--- ![](./evaluation/figures/evalhistory.png) -->

## Klein, NeuroImage 2009

![](./competitions/figures/nihms-130679-f0001.jpg)

* 14 algorithms
* publicly available data
* different labeling protocols
* developers tune their own algorithms

## Klein results

![](./competitions/figures/KleinResults2009.png)

## Take-away message

_"One of the most significant findings of this study is that the relative performances of
the registration methods under comparison appear to be little affected by the choice of
subject population, labeling protocol, and type of overlap measure....  ART, SyN, IRTK, and
SPM's DARTEL Toolbox gave the best results according to overlap and distance measures,
with ART and SyN delivering the most consistently high accuracy across subjects and label sets."_

## EMPIRE 2010

![](./competitions/figures/EMPIRE10criteria.png)

* Alignment of lung boundaries,

* major fissures,

* annotated landmark pairs, and

* topology of displacement field.

## EMPIRE 2010 results

![](./competitions/figures/EMPIRE10Results.png)

## Multi-atlas 2012

## SATA 2013

Three very different problem domains

![](./competitions/figures/SATA_data.png)

## BRATS 2013

![](./competitions/figures/brats2013results1.png)

## BRATS 2013 results

![](./competitions/figures/brats2013results2.png)

[Tustison, et al., Optimal symmetric multimodal templates and concatenated random forests for
supervised brain tumor segmentation (simplified) with _ANTsR_, _Neuroinformatics_.](http://www.ncbi.nlm.nih.gov/pubmed/25433513)

## STACOM 2014

![](./competitions/figures/image_gallery.jpg)

<!--- A great opportunity to showcase a [new SyN variant](http://www.ncbi.nlm.nih.gov/pubmed/24409140) -->

@@

\centering

<BR><BR>

_\`\`Papers are just advertisements for the science.\'\'_
<BR>

@@

NOtes:
* Previously discussed Brians work
* The variant most widely used

@@

     -i, --initialize-transforms-per-stage (1)/0
          Initialize linear transforms from the previous stage. By enabling this option,
          the current linear stage transform is directly intialized from the previous
          stage\'s linear transform; this allows multiple linear stages to be run where
          each stage directly updates the estimated linear transform from the previous
          stage. (e.g. Translation -> Rigid -> Affine).
          <VALUES>: 0

@@

<!--
## Atropos + N4 $\rightarrow$ `antsAtroposN4.sh`

@@

```
-->

<!-- 
## Multi-atlas segmentation

@@

\begin{centering}

\includegraphics[width=3in]{./evaluation/figures/templates.png}

\end{centering}

* Templates representing the average mean shape and intensity are built directly from the cohort to be analyzed, e.g. pediatric vs. middle-aged brains.  
* Acquisition and anonymization (e.g. defacing) protocols are often different.

@@

\begin{centering}

\includegraphics[width=2.5in]{./evaluation/figures/templatePriors.png}

\end{centering}

Each template is [processed](https://github.com/ntustison/antsCookTemplatePriorsExample)
to produce auxiliary images which are used for brain extraction and brain segmentation.

@@

\begin{centering}

\includegraphics[width=3.5in]{./evaluation/figures/brainExtraction.png}

\end{centering}

Comparison with de facto standard FreeSurfer package.  Note the difference in separation of
the gray matter from the surrounding CSF.  (0 failures out of 1205 scans)

@@

\begin{centering}

\includegraphics[width=3.5in]{./evaluation/figures/brainSegmentation.png}

\end{centering}

Randomly selected healthy individuals.  ```Atropos``` gets good performance across ages.

@@

\begin{centering}

\includegraphics[width=4in]{./evaluation/figures/corticalThicknessEstimation.png}

\end{centering}

@@

##  What about using both the T1 & T2?

@@

@@

# Advanced Normalization Tools in R (ANTsR)

## Multimodal Brain Tumor Segmentation (BRATS 2013)

\begin{centering}

\includegraphics[width=4.5in]{./competitions/figures/brats2013results1.png}

\end{centering}

[Tustison, et al., Optimal symmetric multimodal templates and concatenated random forests for
supervised brain tumor segmentation (simplified) with _ANTsR_, _Neuroinformatics_.](http://www.ncbi.nlm.nih.gov/pubmed/25433513)

## White matter hyperintensities in TBI

\begin{centering}

\includegraphics[width=4.5in]{./wmhs/figures/wmhPipeline.png}

\end{centering}

## Social behavior and immunity dysfunction in mice

\begin{centering}

\includegraphics[width=3in]{./antsr/figures/filiano_rsfmri.png}

\end{centering}

## Other ANTsR work

* [Pediatric template of brain perfusion](http://www.nature.com/articles/sdata20153)

* [Automated segmentation of chronic stroke lesions using LINDA: Lesion identification with neighborhood data analysis](http://www.ncbi.nlm.nih.gov/pubmed/26756101)

* [Eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/25448483)

* [ANTsRNet](https://github.com/stnava/ANTsRNet): a growing collection of popular deep learning architectures

## Open source

* https://github.com/stnava/ANTs

* https://github.com/stnava/ANTsR

* https://github.com/stnava/ANTsRNet
